title: "STOTEN_Beta"
# Beta Diversity Calculation and Graphics
#Please download the RMD file and run each chunk separately

#load packages
```{r}
library("rmarkdown")
library(tidyr)
library(tidyverse)
library(dplyr)
library(readxl)
library(ggplot2)
library(pheatmap) 
library(vegan)
library(RColorBrewer)
library(permute)
library(lattice)
library(labdsv)
library(dendextend)
library(ade4)
library(phyloseq)
library("ape")
library("plyr")
library(scales)
library(Hmisc)
library(corrplot)
library(miscTools)
library(gtools)
library("ggpubr")

#R.version.string
```

setwd("/Users/yikeshen/Desktop/Salmonella_Paper")

############## PLEASE RUN plants and soil sample phyloseq separately!######### 
#Processing data, turn into physeq object, plants (Shoots&Roots) only
#Figure3 left panel, Figure S3 left panel

```{r}
rm(list=ls())
setwd("/Users/yikeshen/Desktop/Salmonella_Paper")
OTUtableraw <- read_excel("Salmonella_OTU_Silva_Dec.xlsx", sheet = "OTU_all")
OTUtable <- OTUtableraw %>% filter(!grepl("Archaea",OTUID)) %>% filter(!grepl("Eukaryota",OTUID))
OTU <- OTUtable %>% dplyr::select(-"OTUID")
ROWNAMES <- OTU$FeatureID 
OTU = OTU[,mixedsort(colnames(OTU))]

OTU <- OTU %>% select(-c(S1:S48))#remove soil samples for ordination
OTU <- OTU %>% select(-P1,-P3)#didn't amplify, remove
row.names(OTU) <- OTU$FeatureID 
OTU <- OTU %>% select(-FeatureID)
OTU[is.na(OTU)] <- 0
OTU <- OTU %>% as.matrix()
row.names(OTU) <- ROWNAMES

Taxonomy <- OTUtable %>% 
  dplyr::select("OTUID")

Taxonomy <- Taxonomy %>%
  separate("OTUID",into=c("Domain","Phylum","Class","Order","Family","Genus","Species"),sep=";")

TAX <- Taxonomy %>% as.data.frame() %>% 
  mutate(Domain=gsub("D_0__","",Domain)) %>% 
  mutate(Phylum=gsub("D_1__","",Phylum)) %>% 
  mutate(Class=gsub("D_2__","",Class)) %>% 
  mutate(Order=gsub("D_3__","",Order)) %>% 
  mutate(Family=gsub("D_4__","",Family)) %>% 
  mutate(Genus=gsub("D_5__","",Genus)) %>% 
  mutate(Species=gsub("D_6__","",Species)) %>% 
  as.matrix() %>% as.data.frame()

TAX[is.na(TAX)] <- NA
TAX <- apply(TAX, 2, function(x) gsub("^$|^ $", NA, x))
TAX <- TAX %>% as.matrix()
row.names(TAX) <- ROWNAMES

map <- read_excel("Salmonella_OTU_Silva_Dec.xlsx", sheet = "ENV")
SAMPLEIDROWNAMES <- map$SampleID
map <- map %>% as.matrix() %>% as.data.frame()
row.names(map) <- SAMPLEIDROWNAMES
ROANAMESLISTSAMPLESID <- row.names(map)

OTU=otu_table(OTU,taxa_are_rows = TRUE)
rownames(TAX) <- row.names(OTU)
TAX=tax_table(TAX)
MAP <-  sample_data(map)
physeq = merge_phyloseq(OTU, TAX, MAP)
Tree=rtree(ntaxa(physeq),rooted=TRUE,tip.label = taxa_names(physeq))

physeq = phyloseq(OTU, TAX, MAP,Tree)

```

#Beta Diversity PCoA Weighted Unifrac, plant samples
#Figure 3 left panel
```{r}
theme_set(theme_bw())
physeq.ord_unifrac = ordinate(physeq, "PCoA", "unifrac", weighted=TRUE)
p1 <- plot_ordination(physeq, physeq.ord_unifrac, type="SampleID", color="Antibiotics", shape="Compartment") 
print(p1)
p1+geom_point(size=5, alpha=0.75)+labs(x="PCoA1 [91.9%]", y="PCoA2 [2.7%]")+ggtitle("PCoA on Weighted Unifrac Distance")+theme(plot.title = element_text(hjust = 0.5))+theme(legend.position="right",axis.text=element_text(size=16),axis.title=element_text(size=16),plot.title = element_text(size=22, hjust=0.5),text = element_text(size = 16))+scale_color_manual(breaks = c("NO", "YES"),
                        values=c("blue", "red"))

#slightly different variation explained in first two axes from 2021 validation and 2020 run. Does not influence results
#p value report from Qiime2 run

```

#Beta diversity colored on dates - plant samples
#FigureS3 left panel
```{r}
theme_set(theme_bw())
physeq.ord_unifrac = ordinate(physeq, "PCoA", "unifrac", weighted=TRUE)
p2 <- plot_ordination(physeq, physeq.ord_unifrac, type="SampleID", color="Time", shape="Compartment") 
print(p2)
p2+geom_point(size=5, alpha=0.75)+labs(x="PCoA1 [92.3%]", y="PCoA2 [2.6%]")+ggtitle("PCoA on Weighted Unifrac Distance")+theme(plot.title = element_text(hjust = 0.5))+theme(legend.position="right",axis.text=element_text(size=16),axis.title=element_text(size=16),plot.title = element_text(size=22, hjust=0.5),text = element_text(size = 16))+scale_fill_brewer(palette = "jco")
#slightly different variation explained in first two axes from 2021 validation and 2020 run. Does not influence results
#p value report from Qiime2 run
```


##Processing data, turn into physeq object, beta diversity for soil samples (Rhizosphere and soil samples)
#Figure 3 right panel
#Figure S3 right panel
```{r}
rm(list=ls())
setwd("/Users/yikeshen/Desktop/Salmonella_Paper")
OTUtableraw <- read_excel("Salmonella_OTU_Silva_Dec.xlsx", sheet = "OTU_all")
OTUtable <- OTUtableraw %>% filter(!grepl("Archaea",OTUID)) %>% filter(!grepl("Eukaryota",OTUID))
OTU <- OTUtable %>% dplyr::select(-"OTUID")
ROWNAMES <- OTU$FeatureID 
OTU = OTU[,mixedsort(colnames(OTU))]
OTU <- OTU %>% select(-c(P1:P48))#Remove shoots for ordination
OTU <- OTU %>% select(-S21)#did not amplify, remove
row.names(OTU) <- OTU$FeatureID 
OTU <- OTU %>% select(-FeatureID)
OTU[is.na(OTU)] <- 0
OTU <- OTU %>% as.matrix()
row.names(OTU) <- ROWNAMES

Taxonomy <- OTUtable %>% 
  dplyr::select("OTUID")

Taxonomy <- Taxonomy %>%
  separate("OTUID",into=c("Domain","Phylum","Class","Order","Family","Genus","Species"),sep=";")

TAX <- Taxonomy %>% as.data.frame() %>% 
  mutate(Domain=gsub("D_0__","",Domain)) %>% 
  mutate(Phylum=gsub("D_1__","",Phylum)) %>% 
  mutate(Class=gsub("D_2__","",Class)) %>% 
  mutate(Order=gsub("D_3__","",Order)) %>% 
  mutate(Family=gsub("D_4__","",Family)) %>% 
  mutate(Genus=gsub("D_5__","",Genus)) %>% 
  mutate(Species=gsub("D_6__","",Species)) %>% 
  as.matrix() %>% as.data.frame()

TAX[is.na(TAX)] <- NA
TAX <- apply(TAX, 2, function(x) gsub("^$|^ $", NA, x))
TAX <- TAX %>% as.matrix()
row.names(TAX) <- ROWNAMES

map <- read_excel("Salmonella_OTU_Silva_Dec.xlsx", sheet = "ENV")
SAMPLEIDROWNAMES <- map$SampleID
map <- map %>% as.matrix() %>% as.data.frame()
row.names(map) <- SAMPLEIDROWNAMES
ROANAMESLISTSAMPLESID <- row.names(map)

OTU=otu_table(OTU,taxa_are_rows = TRUE)
rownames(TAX) <- row.names(OTU)
TAX=tax_table(TAX)
MAP <-  sample_data(map)
physeq = merge_phyloseq(OTU, TAX, MAP)
Tree=rtree(ntaxa(physeq),rooted=TRUE,tip.label = taxa_names(physeq))

physeq = phyloseq(OTU, TAX, MAP,Tree)

```

#beta diversity-weighted unifrac distance
```{r}
theme_set(theme_bw())
physeq.ord_unifrac = ordinate(physeq, "PCoA", "unifrac", weighted=TRUE)
p1 <- plot_ordination(physeq, physeq.ord_unifrac, type="SampleID", color="Antibiotics", shape="Compartment") 
print(p1)
p1+geom_point(size=5, alpha=0.75)+labs(x="PCoA1 [34.9%]", y="PCoA2 [21.4%]")+ggtitle("PCoA on Weighted Unifrac Distance")+theme(plot.title = element_text(hjust = 0.5))+theme(legend.position="right",axis.text=element_text(size=16),axis.title=element_text(size=16),plot.title = element_text(size=22, hjust=0.5),text = element_text(size = 16))+scale_color_manual(breaks = c("NO", "YES"),
                        values=c("blue", "red"))

#slightly different variation explained in first two axes from 2021 validation and 2020 run. Does not influence results
#p value report from Qiime2 run
```
#Beta diversity colored on dates - soil samples
#Figure S3, right panel
```{r}
theme_set(theme_bw())
physeq.ord_unifrac = ordinate(physeq, "PCoA", "unifrac", weighted=TRUE)
p2 <- plot_ordination(physeq, physeq.ord_unifrac, type="SampleID", color="Time", shape="Compartment") 
print(p2)
p2+geom_point(size=5, alpha=0.75)+labs(x="PCoA1 [34.9%]", y="PCoA2 [22.3%]")+ggtitle("PCoA on Weighted Unifrac Distance")+theme(plot.title = element_text(hjust = 0.5))+theme(legend.position="right",axis.text=element_text(size=16),axis.title=element_text(size=16),plot.title = element_text(size=22, hjust=0.5),text = element_text(size = 16))+scale_fill_brewer(palette = "jco")
#slightly different variation explained in first two axes from 2021 validation and 2020 run. Does not influence results
#p value report from Qiime2 run

```