#title: "STOTEN_NST"
#This files contains NST calculation and graphics
#Figure 4
#Figure S4

#NST calculation
```{r}
rm(list=ls())
#install.packages("NST")
library(NST)
library(vegan)
setwd("/Users/yikeshen/Desktop/Salmonella_Paper/NST")

prefix="Yike" # define output file prefix.

#Please note, the shoots had too few reads to compute NST, thus shoots samples were removed

####Please run this chunk for Figure4 
#com.file="OTUtable.csv"
com.file="OTULoad.csv"
treat.file="NST_group_compartment.csv"
#####################################

####Please run this chunk for Figure S4
#com.file="OTULoad.csv"
#treat.file="NSTDetailgroup.csv"

nworker=4
rand.time=1000 # set a low number for test. real data should use 1000 times.

# load data
comm=t(read.csv(com.file,header = TRUE,sep=",",row.names = 1))
comm <- t(comm)
#comm <- comm[-c(1:24),]
treat=read.csv(treat.file,header = TRUE,row.names = 1)

dim(comm)
dim(treat)

#check match sample ID
sampc=NST::match.name(rn.list=list(comm=comm,treat=treat))
comm=sampc$comm
dim(comm)
comm=comm[,colSums(comm)>0] # remove OTUs have no read across remained samples.
dim(comm)
treat=sampc$treat

prefixi=paste0(prefix,".bray") # file prefix with beta diversity index name. The abundance-weighted Jaccard is named as Ruzicka.
group=treat[,1,drop=FALSE] # select which column in treatment table will be used
head(group)

t0=Sys.time()

nst1=tNST(comm=comm, group=group, meta.group=NULL, meta.com=NULL,
          dist.method="bray", abundance.weighted=TRUE,
          rand=rand.time, output.rand=TRUE, nworker=nworker,
          LB=FALSE, null.model="PF", between.group=FALSE,
          SES=FALSE, RC=FALSE)
save(nst1,file = paste0(prefixi,".NST.rda"))
write.csv(nst1$index.grp,file=paste0(prefixi,".NST.summary.csv"))
(t1=format(Sys.time()-t0)) # count time cost

# Bootstrapping
t0=Sys.time()
nstbt=nst.boot(nst.result=nst1, group = group, rand = rand.time, trace = TRUE, 
               two.tail = FALSE, out.detail = TRUE, between.group = FALSE, 
               nworker = nworker)
save(nstbt,file = paste0(prefixi,".NST.boot.rda"))
write.csv(nstbt$summary,file=paste0(prefixi,".NST.boot.summary.csv"))
write.csv(nstbt$compare,file=paste0(prefixi,".NST.boot.compare.csv"))
(t2=format(Sys.time()-t0))

# PERMANOVA
t0=Sys.time()
nstpaov=nst.panova(nst.result=nst1, group = group, rand = rand.time, trace = TRUE)
write.csv(nstpaov,file=paste0(prefixi,".NST.PERMANOVA.csv"))
(t3=format(Sys.time()-t0))

#Below are other distant matrix
comm.b=comm
comm.b[comm.b>0]=1
samp.ab=rowSums(comm)
prob.ab=matrix(colSums(comm),nrow=nrow(comm),ncol=ncol(comm),byrow=TRUE)
comm.rand=ab.assign(comm.b,samp.ab,prob.ab)

beta.g(comm, dist.method="bray", abundance.weighted=TRUE,
       as.3col=FALSE,out.list=TRUE)
chaosorensen(comm, dissimilarity=TRUE, to.dist=TRUE)
chaojaccard(comm, dissimilarity=TRUE, to.dist=TRUE)

# calculate one index
beta.bray=beta.g(comm=comm,as.3col=TRUE)
# calculate multiple indexes
beta.td=beta.g(comm=comm,dist.method=c("bray","jaccard","euclidean",
                                       "manhattan","binomial","chao","cao"),
               abundance.weighted = TRUE,out.list=FALSE)

data(beta.limit)

bray=beta.g(comm,dist.method="bray")
bray.3col=dist.3col(bray)

```



#Normalized Stochastic Ratio Graphics

#Figure 4: Normalized stochastic ration of baterial community assembly in the bulk soil, rhizosphere soil, and lettuce roo samples with and without antibiotic exposure
```{r}
#Compartment
rm(list=ls())
library(ggplot2)
temporalraw <- read.csv("Yike.bray.NST.boot.summary_compartment.csv",header = TRUE,row.names = 1)
temporalraw <- temporalraw[,4:8]*100
wrapgroup <- c(rep("Shoots",2),rep("Roots",2),rep("Soil",2),rep("Rhizosphere",2))
Treatment <- c("Control","Antibiotics","Control","Antibiotics","Control","Antibiotics","Control","Antibiotics")
temporalraw <- cbind(temporalraw,rownames(temporalraw),wrapgroup,Treatment)
#Remove shoots for low NST
temporalraw <- temporalraw[-c(1:2),]
ROWNAMESNEW <- cbind("Roots_No","Roots_Yes","Soil_No","Soil_Yes","Rhizo_No","Rhizo_Yes")
rownames(temporalraw) <- ROWNAMESNEW

temporalraw$wrapgroup <- factor(x = temporalraw$wrapgroup,levels = c("Soil","Rhizosphere","Roots"))

pd <- position_dodge(0.1)
ggplot(temporalraw, aes(x=rownames(temporalraw), ymin=Min, ymax=Max, lower=Quantile25, middle=Median, upper=Quantile75,fill=Treatment))+
  geom_boxplot(stat='identity')+xlab("Treatment") +
  ylab("Normalized Stochasticity Ratio (NST) (%)") +
  ggtitle("Stochasticity Test") +
  scale_y_continuous(breaks=0:10*10)+                       # Expand y range
  theme_bw() +scale_fill_manual(breaks = c("Antibiotics", "Control"),values=c("red", "blue"))+
  theme(legend.position="top",axis.text=element_text(size=16),axis.title=element_text(size=16),
        plot.title = element_text(size=22, hjust=0.5),text = element_text(size = 16),
        axis.text.x = element_text(angle = 0))+facet_wrap(~wrapgroup,1,scale="free_x")
```



#Figure S4. Normalized stochastic ratio of samples collected on four sampling days
```{r}
rm(list=ls())
library(scales)
library(ggplot2)
setwd("/Users/yikeshen/Desktop/Salmonella_Paper/NST")
temporalraw <- read.csv("Yike.bray.NST.boot.summary_allgroups.csv",header = TRUE,row.names = 1)
temporalraw <- temporalraw[,4:8]*100
wrapgroup <- c(rep("Shoots",8),rep("Roots",8),rep("Soil",8),rep("Rhizosphere",8))
Treatment <- c(rep("Control",4),rep("Antibiotics",4),rep("Control",4),rep("Antibiotics",4),rep("Control",4),
               rep("Antibiotics",4),rep("Control",4),rep("Antibiotics",4))
temporalraw <- cbind(temporalraw,rownames(temporalraw),wrapgroup,Treatment)
temporalraw <- temporalraw[-c(1:8),]
pd <- position_dodge(0.1)
temporalraw$wrapgroup <- factor(x = temporalraw$wrapgroup,levels = c("Soil","Rhizosphere","Roots"))
ggplot(temporalraw, aes(x=rownames(temporalraw), ymin=Min, ymax=Max, lower=Quantile25, middle=Median, upper=Quantile75, fill=Treatment))+
  geom_boxplot(color="black",stat='identity')+xlab("Harvest Day") +
  ylab("Normalized Stochasticity Ratio (NST) (%)") +
  scale_fill_manual(breaks = c("Antibiotics", "Control"),values=c("red", "blue"))+
  ggtitle("Stochasticity Test") +
  scale_y_continuous(breaks=0:10*10)+                       # Expand y range
  theme_bw() +
  theme(legend.position="top",axis.text=element_text(size=16),axis.title=element_text(size=16),
        plot.title = element_text(size=22, hjust=0.5),text = element_text(size = 16),
        axis.text.x = element_text(angle = 90))+facet_wrap(~wrapgroup,1,scales = "free_x")

```

